\Large \bf 6.829 Final Project: A comparison of bit-rate selection algorithms

### 2013-05-13 paper/paper.tex
our awesome research paper

%don't want date printed
\date{}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Wonderful : A Terrific Application and Fascinating Paper}

%for single author (just remove % characters)
\author{
{\rm Your N.\ Here}\\
Your Institution
\and
{\rm Second Name}\\
Second Institution
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

@@

\subsection*{Abstract}
Your Abstract Text Goes Here.  Just a few facts.
Whet our appetites.

\section{Introduction}

A paragraph of text goes here.  Lots of text.  Plenty of interesting
text. \\

More fascinating text. Features\endnote{Remember to use endnotes, not footnotes!} galore, plethora of promises.\\

\section{This is Another Section}

Some embedded literal typset code might 
look like the following :

{\tt \small
\begin{verbatim}
int wrap_fact(ClientData clientData,
              Tcl_Interp *interp,
              int argc, char *argv[]) {
    int result;
    int arg0;
    if (argc != 2) {
        interp->result = ``wrong # args'';
        return TCL_ERROR;
    }
    arg0 = atoi(argv[1]);
    result = fact(arg0);
    sprintf(interp->result,''%d'',result);
    return TCL_OK;
}
\end{verbatim}
}

Now we're going to cite somebody.  Watch for the cite tag.
Here it comes~\cite{Chaum1981,Diffie1976}.  The tilde character (\~{})
in the source means a non-breaking space.  This way, your reference will
always be attached to the word that preceded it, instead of going to the
next line.

\section{This Section has SubSections}
\subsection{First SubSection}

Here's a typical figure reference.  The figure is centered at the
top of the column.  It's scaled.  It's explicitly placed.  You'll
have to tweak the numbers to get what you want.\\

% you can also use the wonderful epsfig package...
\begin{figure}[t]
\begin{center}
\begin{picture}(300,150)(0,200)
\put(-15,-30){\special{psfile = fig1.ps hscale = 50 vscale = 50}}
\end{picture}\\
\end{center}
\caption{Wonderful Flowchart}
\end{figure}

This text came after the figure, so we'll casually refer to Figure 1
as we go on our merry way.

\subsection{New Subsection}

It can get tricky typesetting Tcl and C code in LaTeX because they share
a lot of mystical feelings about certain magic characters.  You
will have to do a lot of escaping to typeset curly braces and percent
signs, for example, like this:
``The {\tt \%module} directive
sets the name of the initialization function.  This is optional, but is
recommended if building a Tcl 7.5 module.
Everything inside the {\tt \%\{, \%\}}
block is copied directly into the output. allowing the inclusion of
header files and additional C code.'' \\

Sometimes you want to really call attention to a piece of text.  You
can center it in the column like this:
\begin{center}
{\tt \_1008e614\_Vector\_p}
\end{center}
and people will really notice it.\\

\noindent
The noindent at the start of this paragraph makes it clear that it's
a continuation of the preceding text, not a new para in its own right.

Now this is an ingenious way to get a forced space.
{\tt Real~$*$} and {\tt double~$*$} are equivalent.

Now here is another way to call attention to a line of code, but instead
of centering it, we noindent and bold it.\\

\noindent
{\bf \tt size\_t : fread ptr size nobj stream } \\

And here we have made an indented para like a definition tag (dt)
in HTML.  You don't need a surrounding list macro pair.
\begin{itemize}
\item[]  {\tt fread} reads from {\tt stream} into the array {\tt ptr} at
most {\tt nobj} objects of size {\tt size}.   {\tt fread} returns
the number of objects read. 
\end{itemize}
This concludes the definitions tag.

\subsection{How to Build Your Paper}

You have to run {\tt latex} once to prepare your references for
munging.  Then run {\tt bibtex} to build your bibliography metadata.
Then run {\tt latex} twice to ensure all references have been resolved.
If your source file is called {\tt usenixTemplate.tex} and your {\tt
  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
{\tt \small
\begin{verbatim}
latex usenixTemplate
bibtex usenixTemplate
latex usenixTemplate
latex usenixTemplate
\end{verbatim}
}

\subsection{Last SubSection}

@@

\section{Acknowledgments}

A polite author always includes acknowledgments.  Thank everyone,
especially those who funded the work.

\section{Availability}

It's great when this section says that MyWonderfulApp is free software, 
available via anonymous FTP from

\begin{center}
{\tt ftp.site.dom/pub/myname/Wonderful}\\
\end{center}

Also, it's even greater when you can write that information is also 
available on the Wonderful homepage at

\begin{center}
{\tt http://www.site.dom/\~{}myname/SWIG}
\end{center}

Now we get serious and fill in those references.  Remember you will
have to run latex twice on the document in order to resolve those
cite tags you met earlier.  This is where they get resolved.
We've preserved some real ones in addition to the template-speak.
After the bibliography you are DONE.

{\footnotesize \bibliographystyle{acm}
\bibliography{../common/bibliography}}

\theendnotes

>>>>

%don't want date printed
\date{May 2013}

%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf 6.829 Final Project: A comparison of bit-rate selection algorithms}
%for single author (just remove % characters)
\author{
{\rm Colleen Josephson}\\
MIT
\and
{\rm Pavel Panchekha}\\
MIT
% copy the following lines to add more authors
% \and
% {\rm Name}\\
%Name Institution
} % end author

@@

\subsection*{Abstract}
In this paper we compare the performance SampleRate and Minstrel, two popular bit rate selection algorithms that have widespread real-world usage on the Linux operating system. We use a novel trace-based approach that allows us avoid in-depth kernel programming and write simple Python scripts instead. We also introduce improvements to the Minstrel algorithm that allow for significant gains in throughput. 
\section{Introduction}

One of the key ways that wireless networks differ from wired is that wireless networks have varying link rates. Link conditions vary with time due to interference from other devices, changing network geometry, and mobile clients. An optimal rate at some time $t$ may be different from the optimal rate just 30 seconds earlier. A good bit rate selection algorithm has to detect and adapt to these conditions. If the chosen bit rate is too slow, then the throughput will be unnecessarily low. If the rate is too high, then failures will be very frequent and throughput will again suffer.

There are three main classes of bit rate adaptiation protocols: frame-based, SNR-based, and cross-layer protocols. Frame-based protocols measure the fraction of sucessfully received protocols. SNR protocols make decisions based on the estimated Signal to Noise ratio. Cross-layer protocols use SoftPHY data from the physical layer. The most commonly implemented protocols on today's networks are frame based, because SNR protocols perform poorly [CITATION HERE], and cross-layer protocols cannot be deployed on current networks because they violate the network layering abstration.

The two most popular fram-based protocols are SampleRate and Minstrel. Both were implemented in the MadWifi drivers for Linux wirless, and today Minstrel is the default bit-rate selection algorithm for all wireless drivers on Linux. Our research used a trace-based approach to analyze the performance of these two algorithms. We also created a modified version of Minstrel that provides significant throughput gains over the vanilla Minstrel implementation.

This paper will provide a general overview of SampleRate and Minstrel. We then discuss our testing methodology. We compare the performance of the two algorithms, and then discuss modifications we made to Minstrel and compare it's performance to the current kernel implementation.
\section{SampleRate}

Sample all the rates~\cite{samplerate, sprout, minstrel}.

\section{Minstrel}

Derek Smithies, Ph-fucking-D.

\section{Methodology}

We did some things using some methods.

\subsection{Motivation}

MadWifi is old and sucks plus we'd need to buy hardware off e-bay. The current Atheros drivers use the linux default, Minstrel. There is no kernel implementation of SampleRate and porting it woile have been extremely annoying (possibly impossible) given the time constraints. Etc. Etc. Blah Blah.

\subsection{Trace Collection}

Yeee haw.

\subsection{Testing framework}

Harness.py, etc.

\section{Analysis}

@@

\subsection{Improvements to Minstrel}
Minproved.

\section{Future Work}

802.11n, SampleRate with EWMA, take data using broadcast approach,  submitting minstrel changes as kernel patch?

\section{Conclusion}

The end.

{\footnotesize \bibliographystyle{acm}
\bibliography{paper}}

\section{Acknowledgements}
Jonathan Perry, Hari Balikrishnan, and Derek Smithies.
